# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
fitdata<-read.csv("activity.csv")
fitdata[,2]<-as.Date(fitdata[,2], format="%Y-%m-%d")
```


## What is mean total number of steps taken per day?
- The histogram  

```{r}
plot(fitdata[,2], fitdata[,1], type="h", lwd="5")
```

- The mean steps taken per day  

```{r}
sapply(split(fitdata[,1], fitdata[,2]), mean)
```

          
- The median steps taken per day  

```{r}
sapply(split(fitdata[,1], fitdata[,2]), median)
```


## What is the average daily activity pattern?

Time series plot of 5-minute interval vs average number of steps taken

```{r}
fitdata<-fitdata[complete.cases(fitdata),]
plot(unique(fitdata[,3]), sapply(split(fitdata[,1], fitdata[,3]), mean), type="l", ylim=c(0,250))
```

5-minute interval having the highest average steps taken per day

```{r}
temp=subset(fitdata, sapply(split(fitdata[,1], fitdata[,3]), mean)==max(sapply(split(fitdata[,1], fitdata[,3]), mean)))
unique(temp[,3])
```
## Imputing missing values

Reloading the original file first containing NA values
```{r}
fitdata<-read.csv("activity.csv")
fitdata[,2]<-as.Date(fitdata[,2], format="%Y-%m-%d")
```

Then, find the number of rows containing missing values
```{r}
temp<-fitdata[complete.cases(fitdata),]
length(fitdata[,1])-length(temp[,1])
```

Creation of the new dataset. Logic is that the missing values
are replaced by the average of the number of steps taken over the entire two months

```{r}
fitdata[is.na(fitdata)]=floor(mean(fitdata[,1], na.rm=TRUE))
```

The histogram, means and medians with missing values replaced

```{r}
plot(fitdata[,2], fitdata[,1], type="h", lwd="5")
sapply(split(fitdata[,1], fitdata[,2]), mean)
sapply(split(fitdata[,1], fitdata[,2]), median)
```

Because of using the logic given above, not much changes
in the dataset except that the mean and median of the days with NA values
previously increases

## Are there differences in activity patterns between weekdays and weekends?

Let's first create the two separate subsets
```{r}
fitdata$day<-weekdays(fitdata[,2])
wkds<-subset(fitdata, fitdata[,4]!="Saturday" & fitdata[,4]!="Sunday")
wkes<-subset(fitdata, fitdata[,4]=="Saturday" | fitdata[,4]=="Sunday")
```

Then, we plot the timeseries with the new datasets(after setting the mfrow appropriately)
```{r}
par(mfrow=c(2,1))
plot(unique(wkds[,3]), sapply(split(wkds[,1], wkds[,3]), mean), type="l", ylim=c(0,200))
plot(unique(wkes[,3]), sapply(split(wkes[,1], wkes[,3]), mean), type="l", ylim=c(0,200))
```
